---
title: "BCO6007 Leture 4"
author: "MariaProkofieva"
date: "01/04/2020"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, echo=FALSE, warning=FALSE}

knitr::opts_chunk$set(echo = TRUE,   
                      message = FALSE,
                      warning = FALSE,
                      collapse = TRUE,
                      fig.height = 4,
                      fig.width = 8,
                      fig.align = "center",
                      cache = FALSE,
                      comment = NA )

library(tidyverse)
library(ggplot2)

```

# Data Preparation and transformation

## Objectives

1. understand the concept of tidy data
2. being able to prepare data for further analysis using `tidyverse`
3. being able to use visualizations to check "preparation" of data
4. use data wrangling approaches to transform the data to fit the needs of data analysis

----------------

*The examples are based on using `tidyverse` library, so you need to install it first if you have not! Uncomment the next `r` chunk and run it in your console*

`install.packages("tidyverse")`

----------------

Let's load `tidyverse` library and load our data set


```{r readgrad}
library(tidyverse)
grad <- read_csv("data/graduate-programs.csv")
grad %>% top_n(10)
```

Let's count:

```{r grad-count}
head (grad)
grad %>% count(subject)
```

Let's filter

```{r grad-filter}
grad %>%
  filter(subject == "economics") %>%
  summarise(meanStudents = mean(NumStud),
          s = sd(NumStud))
```

Let's make a plot

```{r grad-plot}
grad %>%
  filter(subject == "economics") %>%
  ggplot(aes(x = NumStud, 
             y = MedianTimetoDegree)) +
  geom_point() + 
  theme(aspect.ratio = 1)
```


## Tidy data

Let's have a look at the data....

```{r}
table1

table2

table3

table4a

table4b
```

Which one is easier to use? E.g. for visualization:

```{r}
# Visualise changes over time
library(ggplot2)
ggplot(table1, aes(year, cases)) + 
  geom_line(aes(group = country), colour = "grey50") + 
  geom_point(aes(colour = country))
```

Let's tidy our dataset table4a

```{r}

table4a

table4a %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")
```

Now, it's your turn to "tidy" `table4b`

``` {r}
table4b

table4b %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "population")
```

Now, let's have a look at `table2`

```{r}
table2

table2 %>%
    pivot_wider(names_from = type, values_from = count)
```