---
title: "BCO6007: Lecture 5"
subtitle: "Dealing with different datatypes"
author: "Dr Maria Prokofieva"
institute: "VU Business School"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      ratio: "16:9"

      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = TRUE,   
                      message = FALSE,
                      warning = FALSE,
                      fig.height = 4,
                      fig.width = 8,
                      fig.align = "center")
library(tidyverse)
library(lubridate)
library(maps)
```
#Key points

- `tidyverse` package
- `maps` package
- data types
- data viz for different data types

.pull-left[
##Basic functions to review

- tibbles
- factors
- date/times
- maps
]
.pull-right[
If you want to get quick info on any function you type in the console
`?NameOfFunction`
e.g 
```{r}
?slice
```
]
---
#Workflow for the lecture

1. Open your `RStudio.cloud` `BCO6007` project

2. Create a new lecture R script document - e.g. lecture5.R

3. Practice all lecture examples there

4. If your datasets from lecture 3 (lecture 4) are not there, load them using 

---

# Tibbles and tidy data

`Tibbles` are data frames that satisfy the tidy data approach!
You can access individual variables in your tibble by using their position:

# Extract by name or by number!

```{r}
grad<-read_csv("graduate-programs.csv")
avocado<-read_csv("avocado.csv")

grad$subject
grad[, 1]
```

We can also access individual rows! using their number
```{r}
grad[5,]

#OR

grad[5,8]
```
---
# Factors

In R, `factors` are used to work with *categorical variables*, variables that have a fixed and known set of possible values. They are also useful when you want to display character vectors in a non-alphabetical order.

Can you give me examples of categorical variables?

Have a look at the `avocado` dataset and tell which variables "should" be categorical ones

---
# Working with factors

If for some reason, your data is recorded as a character string and you need to convert it to a factor, use `as_factor()` command.

Have a look at the `Help` and tell me how to do this for `region` variable in the `avocado` dataset

Also, let's have a look at our `grad` dataset and identify "possible" factors there

---
# Modifying factor order

It’s often useful to change the order of the factor levels in a visualisation. 

```{r}
grad_summary <- grad %>%
  group_by(Inst) %>%
  summarise(
    students = mean(NumStud, na.rm = TRUE),
    faculty = mean(TotFac, na.rm = TRUE),
    n = n()
  )

grad_summary %>%
  slice_max(students, n=10)%>%
ggplot(aes(students, Inst)) + 
  geom_point()
```
---
It is difficult to interpret this plot because there’s no overall pattern. We can improve it by reordering the levels of `Inst` using `fct_reorder()`. 

`fct_reorder()` takes three arguments:

- `f`, the factor whose levels you want to modify.

- `x`, a numeric vector that you want to use to reorder the levels.

- Optionally, `fun`, a `function` that’s used `if` there are multiple values of `x` for each value of `f`. 
---

Let's change it 
```{r}
grad_summary %>%
  slice_max(students, n=10)%>%
ggplot(aes(students, fct_reorder(Inst, students))) + 
  geom_point()
```
---
Task:

With `avocado` data, 

1. create a summary dataset by `summarise`'ing by the `region` to calculate average `AveragePrice` 

---

```{r}
avocado_summary <- avocado %>%
  group_by(region) %>%
  summarise(
    pricePerRegion = mean(AveragePrice, na.rm = TRUE),
    n = n()
  )
```
---
Let's use `slice_max` and `fct_reorder` to make a `geom_point` plot to present changes in the `pricePerRegion` for the most expensive 15 regions
---

Let's change it 
```{r}
avocado_summary %>%
  slice_max(pricePerRegion, n=15)%>%
ggplot(aes(pricePerRegion, fct_reorder(region, pricePerRegion))) + 
  geom_point()
```
---
# fct_lump

Sometimes you have TOO MANY factors and want to "lump them" together into  the small groups to make a plot or dataset simpler. That’s `fct_lump()`:


By default it lump together the smallest groups 
```{r}
grad_lump<-grad%>%
  mutate(subject = fct_lump(subject)) %>%
  count(subject)

grad_lump
```
---

---
# Working with different data types and evaluating the quality of the data

`skimr` package is fantastic to give "sense" to your dataset. Make sure you `install.packages("skimr")` before the first use!

```{r}
library(skimr)

skim(avocado)
```
---
Let's look at `grad` data and work in pairs to discuss

```{r}
skim(grad)
```
---
# Dates and times

- Does every year have 365 days?
- Does every day have 24 hours?
- Does every minute have 60 seconds?

Working with different date/time formats can be a nightmare or can bea breeze!
with library `lubridate`
---
Make sure to install the library first

```{r eval=FALSE}
install.packages("lubridate")
```
and then load it
```{r}
library(lubridate)
```
---

Let's have some fun with some basic functions there

``` {r}
today()
now()
```
---
Have a look at the examples below and create your own!

```{r}
ymd("2017-01-31")
ymd(20170131)

dmy("31-Jan-2017")

ymd_hms("2017-01-31 20:11:59")
mdy_hm("01/31/2017 08:01")
```
---
# Time spans

- **durations**, which represent an exact number of seconds.
- **periods**, which represent human units like weeks and months.
- **intervals**, which represent a starting and ending point.
---
# Durations

when you subtract two dates, you get a difftime object:

How old is you?
```{r}
my_age <- today() - ymd(19000112)
my_age
```
---
A `difftime` class object records a time span of seconds, minutes, hours, days, or weeks. This ambiguity can make `difftimes` a little painful to work with, so `lubridate` provides an alternative which always uses seconds: the duration.

```{r}
dseconds(15)

# Also
dseconds(15)
dminutes(10)
dhours(c(12, 24))
ddays(0:5)
dweeks(3)
dyears(1)
```
---
# Strings = At the beginning there was a word!

### Hello world! ;)

You can create strings with either single quotes or double quotes.

```{r}
string1 <- "This is a string"
string2 <- 'If I want to include a "quote" inside a string, I use single quotes'
```
---
# String length

`str_length()` tells you the number of characters in a string

```{r}
head(str_length(grad_summary$Inst))
```
---
# Combining strings

To combine two or more strings, use `str_c()`.

Use the `sep` argument to control how they’re separated:

```{r}
head(
  str_c(grad$subject, grad$Inst, sep=" ") 
)
```

---
# Subsetting strings

You can extract parts of a string using `str_sub()`. 

`str_sub()` takes start and end arguments which give the (inclusive) position of the substring:

```{r}
head(
  str_sub(grad$subject, start=1, end=3)
)

# negative numbers count backwards from end
head(
  str_sub(grad$Inst, start=-3, end=-1)
)
```

---
# Maps

To use `maps library`

The maps package contains world mapping data as well as routines for displaying maps of countries, regions etc. The data are stored and accessed via a simple database. High resolution versions of these databases are provided by an additional package (mapdata) and compatible routines to convert between different mapping projections are provided by the mapproj package.
- maps


```{r}
library(maps)
library(mapdata)
library(sp)
aus<-map("worldHires", "Australia", fill=TRUE, xlim=c(110,160),
  ylim=c(-45,-5), mar=c(0,0,0,0))
```





