---
title: "BCO6007: Lecture 5"
subtitle: "Dealing with different datatypes"
author: "Dr Maria Prokofieva"
institute: "VU Business School"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      ratio: "16:9"

      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = TRUE,   
                      message = FALSE,
                      warning = FALSE,
                      fig.height = 4,
                      fig.width = 8,
                      fig.align = "center")
library(tidyverse)
```
#Key points

- `tidyverse` package
- data types
- data viz for different data types

.pull-left[
##Basic functions to review

- tibbles
- factors
- date/times
- time series basi
]
.pull-right[
If you want to get quick info on any function you type in the console
`?NameOfFunction`
e.g 
```{r}
?slice
```
]
---
#Workflow for the lecture

1. Open your `RStudio.cloud` `BCO6007` project

2. Create a new lecture R script document - e.g. lecture5.R

3. Practice all lecture examples there

4. If your datasets from lecture 3 (lecture 4) are not there, load them using 



